---
title: Integrated farming and global warming mitigation in Brazil
author: 
  - name: Hugo Tameirão Seixas
    email: seixas.hugo@protonmail.com
    url: hugotseixas.com
    affiliations:
      - name: State University of Campinas
toc: true
toc-title: Contents
format: 
  html:
    theme: lux
page-layout: article
bibliography: references.bib
link-citations: true
abstract: "There is a high expectation in the reduction of liquid greenhouse gases in Brazil by the implementation of integrated crop-livestock-forestry systems. The foundation of such expectation relies in the hypothesis that these systems are capable of storing more above and below ground carbon, in relation to conventional farming systems, due to the presence of trees. However, that are not robust estimates about the potential of these systems to store carbon in a national scale, and usually point estimates are used to extrapolate carbon sequestration. In order to improve the representation of carbon sequestration by integrated crop-livestock-forestry systems, this project aims to use published data to develop a spatially explicit model to estimate carbon fluxes in Brazil."
appendix-style: plain
license: CC BY-NC-ND 4.0
---

```{r}
#| include: false

library(DiagrammeR)

```

------------------------------------------------------------------------

# Introduction

The implementation of integrated crop-livestock-forestry (ICLF) systems in Brazil are expected to sequester enough carbon to contribute to reduce annual liquid carbon emissions. Despite the expectations, there are not clear estimates of the magnitude of reductions of liquid emissions caused by ICLF expansion. Field studies show contrasting results. @Valani2022 observed a higher organic carbon content in the soil in relation to conventional grazing and other integrated systems. @Bieluczyk2020 field measurements did not show clear differences between extensive grazing and ICLF soil organic matter, with highly temporal heterogeneity in observations. @Silva2018837 found inconsistent differences of soil total organic carbon between pasture and ILPF system due to high heterogeneity of observations along the soil layers, and observed small differences between agriculture and integrated system, in which the agriculture presented higher carbon stocks. @Oliveira2018105 found higher amounts of soil organic carbon of two ICLF systems and continuous pasture areas, there were also observed differences between the two integrated systems.

Unfortunately, estimates of carbon sequestration by ICLF at larger scales usually use point estimates, which can severely bias the results towards one direction. @Azevedo_2018 used a CO<sub>2</sub>-eq removal factor of -6.24 t CO<sub>2</sub>-eq ha<sup>-1</sup> y<sup>-1</sup>, which was extracted from one experiment comparing integrated crop-livestock (ICL) systems to pasture [@Carvalho_2010], and did not observe consistent differences between these types of systems. The removal factor was then extrapolated over estimated ILPF area, this simplification is the best guess that Brazil have over the potential of integrated systems in reducing liquid emissions. In face of such scenario, this study aims to generate estimations of CO<sub>2</sub>-eq removal accounting for variation in many aspects, soil attributes, ICLF system characteristics, species used, spacing and climate. We expect that such estimations will provide more realistic ranges of possible outcomes in a diverse set of future scenarios of ICLF expansion in Brazil.

------------------------------------------------------------------------

# Objectives

The present work is composed by three main objectives, accompanied by parallel related questions: 

1. Identify ICLF systems characteristics and dynamics in Brazil;   
    
    1.1. What are the management characteristics of ICLF systems?
    
    1.2. What are the compositions of each component of ICLF systems?
    
    1.3. What is the productivity of each component of ICLF systems? 
    
    1.4 What are the effects of ICLF system over soil properties?
    
    1.5. What is the economical revenue of ICLF systems?
    
    1.6. What is the potential expansion of ICLF systems in Brazil?

2. Estimate the potential of ICLF systems to provide benefits, in the context of climate change;

    2.1. What are the ICLF system management characteristics that provide the best balance between carbon removal, productivity and financial revenue?
    
    2.2. What are the main factors that influence the carbon removal in the modelling framework?

    2.3. What is the pottential of the expansion of ICLF systems over the emissions reduction pledges made by Brazil?

3.  Evaluate the implementation of ICLF as a mean to curb and mitigate climate change, in relation to different perspectives of solutions to avoid climatic catastrophe;

    3.1. Does the expansion of ICLF systems taking into account equity and possible social impacts, as a public policy?
    
    3.2. The actual strategies to promote the adoption of ICLF systems are being effective?

------------------------------------------------------------------------

# Materials and Methods

In order to achieve the proposed objective, the work will be divided in three major parts:

1. Overview of ICLF characteristics in Brazil;

2. Future projections and estimates of benefits generated by ICLF systems adoption;

3. Critical analysis of ICLF as a public policy to curb climate change;

```{r}
#| echo: false

grViz(
  "digraph {
  
  graph [layout = dot, rankdir = TD, nodesep = 0.9, ranksep = 0.3]
  
  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.5,
        fontsize = 12]

  edge [penwidth = 1.5]

  a [label = 'Systematic \n Review', xlabel = '1']
  b [label = 'Experts Survey', xlabel = '1']
  c [label = 'ICLF overview \n data set', xlabel = '1']
  d [label = 'Projections \n Scenarios', xlabel = '2']
  e [label = 'LULC \n Projections', xlabel = '2']
  f [label = 'Benefits \n Estimations', xlabel = '2']
  g [label = 'Targeted \n Review', xlabel = '3']
  h [label = 'Narrative \n Discussion', xlabel = '3']
  
  a -> c b -> c
  c -> d d -> e
  e -> f c -> f
  f -> h g -> h
  
  }"
)

```

### Integrated crop-livestock-forest overview

#### Literature review

#### Experts survey

### Model projections

```{r}
#| include: false

grViz(
  "digraph {
  
  graph [layout = dot, rankdir = LR, nodesep = 0.9, ranksep = 0.3]
  
  edge [penwidth = 1.5]
  
  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.9,
        fontsize = 12]
        
  a [label = 'Degraded \n Pastures']
  b [label = 'Temperature']
  c [label = 'Precipitation']
  d [label = 'Potential \n Evapotranspiration']
  e [label = 'Soil Texture']
  f [label = 'System \n Management']
  g [label = 'LULC/Management \n Time Series']
  h [label = 'Productivity']
  i [label = 'Litter']
  j [label = 'Decomposition Rates']
  k [label = 'Below Ground Carbon']
  l [label = 'Above Ground Carbon']
  m [label = 'Emissions']
  n [label = 'Carbon Balance']
  
    node [shape = ellipse,
        fontname = Helvetica,
        style = dashed,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.5,
        fontsize = 12]
        
  1 [label = 'LULC Model']
  2 [label = 'Soil Carbon Model']
  3 [label = 'Productivity Model']
  4 [label = 'Economic Model']
  
  a -> 1 f -> 1 1 -> g
  g -> 2 b -> 2 c -> 2 d -> 2 e -> 2 i -> 2 j -> 2 2 -> k h -> 2
  g -> 3 b -> 3 c -> 3 e -> 3 3-> h 3 -> i h -> i 2 -> l i -> l h -> m
  h -> 4
  k -> n l -> n m -> n
  
  }"
)

```

#### Projections scenarios

#### LULC projections

```{r}
#| echo: false

grViz(
  "digraph {
  
  graph [layout = dot, rankdir = LR, nodesep = 0.9, ranksep = 0.3]
  
  edge [penwidth = 1.5]
  
  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.9,
        fontsize = 12]
        
  a [label = 'Degraded \n Pastures']
  b [label = 'System \n Management']
  c [label = 'LULC Scenarios']
  d [label = 'LULC/Management \n Time Series']
  
    node [shape = ellipse,
        fontname = Helvetica,
        style = dashed,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.5,
        fontsize = 12]
        
  1 [label = 'LULC Model']
  
  a -> 1 b -> 1 c -> 1 1 -> d
  
  }"
)

```

#### Benefits estimations

```{r}
#| echo: false

grViz(
  "digraph {
  
  graph [layout = dot, rankdir = LR, nodesep = 0.9, ranksep = 0.3]
  
  edge [penwidth = 1.5]
  
  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.9,
        fontsize = 12]
        
  a [label = 'LULC/Management \n Time Series']
  b [label = 'Temperature']
  c [label = 'Precipitation']
  d [label = 'Potential \n Evapotranspiration']
  e [label = 'Soil Texture']
  h [label = 'Productivity']
  i [label = 'Litter']
  j [label = 'Decomposition Rates']
  
  node [shape = ellipse,
        fontname = Helvetica,
        style = dashed,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 1.5,
        width = 1.5,
        fontsize = 12]
        
  1 [label = 'Productivity Model']
  2 [label = 'Soil Carbon Model']
  3 [label = 'Economic Model']
  
  a -> 1 a -> 2 a -> 3
  
  b -> 1 c -> 1 d -> 1 e -> 1
  1 -> h 1 -> i
  
  b -> 2 c -> 2 d -> 2 e -> 2 j -> 2 i -> 2
  
  h -> 3
  
  }"
)

```

### Critical analysis

------------------------------------------------------------------------

# Results

------------------------------------------------------------------------

# Discussion

------------------------------------------------------------------------

# Conclusions

------------------------------------------------------------------------

# Supporting Information

# References
