---
title: "Integrated farming and global warming mitigation in Brazil"
author: 
  - name: Hugo Tameirão Seixas
    url: https://hugotseixas.com/
    email: https://hugotseixas.com/
  - name: David Montenegro Lapola
institute: UNICAMP
format: 
  revealjs:
    footer: "[github.com/hugotseixas/iclfs-brazil](https://github.com/hugotseixas/iclfs-brazil)"
    transition: fade
    theme: [default, custom.scss]
    reference-location: document
---

```{r}
#| include: false

library(DiagrammeR)

```

## Why?

::: {.sectionhead .absolute top="0" left="90%"}
1 [2 3]{style="opacity:0.25"}
:::

<br>
<br>

::: {.incremental}

To know the potential of Integrated Livestock-Crop-Forestry (ICLF) systems in Brazil

: Reduce national greenhouse gases liquid emissions?

: Increase overall productivity?

: Improve farm revenue?

: Lead agriculture towards sustainability?

:::

## How?

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
[<br>]{.mbr}
[<br>]{.mbr}

::: {.incremental}

1. Identify ICLF systems characteristics and dynamics in Brazil

2. Estimate the potential of ICLF systems to provide benefits, in the context of climate change

3. Evaluate the implementation of ICLF as a mean to curb and mitigate climate change, in relation to different perspectives of solutions

:::

---

### ICLF characterization

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
<br>
[<br>]{.mbr}

Systematic review and meta-analysis:
: Documents from Scopus and Embrapa
: Extraction of experiments data

[<br>]{.mbr}

Experts survey:
: Online questionnaire
: Fill gaps from literature

---

### ICLF characterization

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
[<br>]{.mbr}


- Management characteristics of ICLF systems
    
- Compositions of each component of ICLF systems
    
- Production of each component of ICLF systems
    
- Effects of ICLF system over soil properties
    
- Economical revenue of ICLF systems
    
- Potential expansion of ICLF systems in Brazil

---

### Model projections

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

[<br>]{.mbr}
[<br>]{.mbr}

::: {.customlist}

Projection scenarios:
: ICLF expansion trajectory
: ICLF management

[<br>]{.mbr}

LULC projections:
: Random process
: Not spatial

[<br>]{.mbr}

Benefits estimations:
: Based on empirical and mechanistic processes
: Uses deterministic and probabilistic factors

:::

---

### Model projections

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
<br>
<br>

- Accumulated carbon emission and removal
    
- Food and fiber productivity
    
- Total revenue

---

### Critial analysis

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
<br>
[<br>]{.mbr}

Targeted review:
: Books and institutional documents
: Media news

[<br>]{.mbr}

Results from data:
: Contextualized numbers
: Expectations for the future

---

### Critial analysis

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

<br>
<br>
<br>

- Social impacts of the expansion of ICLF
    
- Relation with other solutions to curb climate change
    
- Long term consistence

---

### Overview

::: {.sectionhead .absolute top="0" left="90%"}
[1]{style="opacity:0.25"} 2 [3]{style="opacity:0.25"}
:::

[<br>]{.mbr}

```{r}
#| echo: false

grViz(height = 600,
  "digraph {
  
  graph [layout = dot,
         rankdir = TD, 
         nodesep = 4, 
         ranksep = 0.4,
         tooltip = ' ']
  
  node [shape = box,
        fontname = Helvetica,
        style = filled,
        fontcolor = '#343a40'
        fillcolor = '#f7f7f9'
        penwidth = 2.5,
        width = 3,
        height = 1.2,
        fontsize = 25,
        tooltip = ' ']

  edge [penwidth = 2.5, 
        arrowsize = 2,
        tooltip = ' ']

  a [label = 'Systematic \n Review']
  b [label = 'Experts Survey']
  c [label = 'ICLF overview \n data set']
  d [label = 'Projections \n Scenarios']
  e [label = 'LULC \n Projections']
  f [label = 'Benefits \n Estimations']
  g [label = 'Targeted \n Review']
  h [label = 'Narrative \n Discussion']
  
  a -> c b -> c
  c -> d d -> e
  e -> f c -> f
  f -> h g -> h
  
  }"
)

```

## What?

::: {.sectionhead .absolute top="0" left="90%"}
[1 2]{style="opacity:0.25"} 3
:::

<br>
<br>

- A database with data from ICLF experiments

- A model for quantifying benefits provided by ICLF systems

- Data set with future projections regarding the expansion of ICLF

- Qualitative assessment of the ICLF as public policy

---

### Expected results

::: {.sectionhead .absolute top="0" left="90%"}
[1 2]{style="opacity:0.25"} 3
:::

<br>
<br>
<br>

- Broad view of potential outcomes from the expansion of ICLF

- Quantitative and qualitative characterization of ICLF systems in Brazil

- Advance in methods to evaluate agricultural systems

---

### Principles

::: {.sectionhead .absolute top="0" left="90%"}
[1 2]{style="opacity:0.25"} 3
:::

[<br>]{.mbr}

Reproducibility
  : Code version control  
  : Detailed methods description
  : Data management plan
    
[<br>]{.mbr}
    
Transparency
  : Code, data and documents freely accessible (with license)

[<br>]{.mbr}
   
Accessibility
  : Accessible document formats
  : Data visualization
  : Simple language

---

## Links

<br>
<br>

lab: <https://labterra.cpa.unicamp.br/>

<br>

website: <https://hugotseixas.com/>

<br>

email: <seixas.hugo@protonmail.com>
